---
title: "Assignment 3 - html"
format: html
editor: visual
echo: false
---

## Assignment 3

#### Task 1

```{r, warning=FALSE, message = FALSE}

rm(list=ls())
library(jsonlite)
library(tidyverse)

df <- fromJSON("https://static01.nyt.com/newsgraphics/2021/12/20/us-coronavirus-deaths-2021/ff0adde21623e111d8ce103fedecf7ffc7906264/scatter.json") %>% 
  mutate(state = state.abb[match(name, state.name)])

df[9, 7] = "D.C"
```

```{r, warning=FALSE, message = FALSE, fig.subcap="Source: New York Times database of reports from state and local health agencies, Centers for Disease Control and Prevention Notes: Chart shows deaths reported since universal vaccine eligibility on April 19, 2021. Vaccination rate is the percentage of the total population fully vaccinated as of Dec. 20, 2021."}

# Making a plot in p1
p1 <- ggplot(df, aes(x = fully_vaccinated_pct_of_pop, 
               y = deaths_per_100k, 
               label = state)
       ) +
  
  geom_point(color = "forestgreen", alpha = 0.5, 
             size = 2, stroke = 0.5
             ) +
  
  geom_text(nudge_y = 0.4, size = 2.5, color = "grey50"
            ) +
  
  geom_segment(x = 0.545, y = 18.2, 
               xend = 0.52, yend = 19, 
               arrow = arrow(length = unit(0.2, "cm"), angle = 15)
               ) +
  
  annotate("text", x=0.56, y=17.5, size=2, face = "bold",
          label= "Lower vaccination rate, \n higher death rate"
          ) +
  
  geom_segment(x = 0.73, y = 9.5, 
               xend = 0.76, yend = 8.6, 
               arrow = arrow(length = unit(0.2, "cm"), angle = 18) 
               ) +
  
  annotate("text", x=0.74, y=10.5, size=2, face = "bold",
          label= "Higher vaccination rate, \n lower death rate"
          ) +
  
  labs(x = "Share of population fully vaccinated",
       y = NULL,
       subtitle = "avg. monthly deaths per 100.000",
       title = "Covid-19 deaths since universal adult vaccine eligibility compared with \n vaccination rates"
       ) +
  
  scale_y_continuous(labels = function(x) ifelse(x==0, "", x), #removing 0 label from y-axis
                     limits = c(0, 20), #setting limits of y-axis
                     expand = c(0, 0) #to center the figure
                     ) +
  
  scale_x_continuous(labels = scales::percent_format(), #changing x-axis label to percentage
                     limits = c(0.45, 0.8), #setting limits of x-axis
                     breaks = seq(0, 1, 0.05), #setting interval of x-axis label(every 5%)
                     expand = c(0, 0)
                     ) +
  theme_bw() + #adding white background
  
  #adding some touches to make it more alike the original plot
    theme(plot.subtitle=element_text(size = 7, color = "grey50", vjust = -1, hjust = 0.01),
          plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
          axis.text.y = element_text(vjust = -0.7, hjust = 1.5, size = 7, color = "grey50"),
          axis.text.x = element_text(color = "grey50"),
          axis.title.x = element_text(color = "grey50", size = 7),
          panel.border = element_rect(color = "grey50", fill = NA, size = 0.5, linetype = "dotted")
          )
#printing
p1
```

#### Task 2

```{r}
lm(deaths_per_100k ~ fully_vaccinated_pct_of_pop, data = df)
```

-36.66 is the value of the coeffiecient. It´s an estimated slope. Since we are working with percentage that would mean that for every percent change in vaccination, the death rate per 100k would decrease by approximately 0.37.

The intercept value is the y-value where x would be equal to 0. It´s an estimate of what the death rate for covid would be per 100k if nobody were vaccinated.

```{r, warning=FALSE, message = FALSE, fig.subcap="Source: New York Times database of reports from state and local health agencies, Centers for Disease Control and Prevention Notes: Chart shows deaths reported since universal vaccine eligibility on April 19, 2021. Vaccination rate is the percentage of the total population fully vaccinated as of Dec. 20, 2021."}
p1 + geom_smooth(method = lm, se = FALSE, size = 0.7)
```
